# Supplemental Code {#sec-code}

## High Dimensional Mediation Analysis Code

```{r read-in_hima, include=FALSE}
knitr::read_chunk("project_setup/functions/hima_functions.R")
knitr::read_chunk("project_setup/functions/hima_plot_functions.R")
```


### Early Integration of omics datasets {#sec-code_hima_early}

#### Run Analysis
```{r HIMA_early, eval=FALSE}
```
#### Plot
```{r HIMA_early_plot, eval = FALSE}
```

### Intermediate Integration of omics datasets {#sec-code_hima_int}
#### Run Analysis
```{r HIMA_intermediate, eval=FALSE}
```
#### Plot
```{r HIMA_intermediate_plot, eval=FALSE}
```

### Late Integration of omics datasets {#sec-code_hima_late}
#### Run Analysis
```{r HIMA_late, eval=FALSE}
```
#### Plot
```{r HIMA_late_plot, eval=FALSE}
```

## Mediation with Latent Factors Analysis Code

```{r read-in_mediation_with_latent_fctrs, include=FALSE}
knitr::read_chunk("project_setup/functions/mediation_with_latent_fctrs_function.R")
knitr::read_chunk("project_setup/functions/mediation_with_latent_fctrs_plot_function.R")
```

### Early Integration of omics datasets {#sec-code_med_with_latent_fctrs_early}

#### Run Analysis
```{r Mediation with latent fctrs early, eval=FALSE}
```
#### Plot
```{r Mediation with latent fctrs early plot, eval=FALSE}
```

### Intermediate Integration of omics datasets {#sec-code_med_with_latent_fctrs_int}
#### Run Analysis
```{r Mediation with latent fctrs intermediate, eval=FALSE}
```
#### Plot
```{r Mediation with latent fctrs intermediate plot, eval=FALSE}
```

### Late Integration of omics datasets {#sec-code_med_with_latent_fctrs_late}
#### Run Analysis
```{r Mediation with latent fctrs late, eval=FALSE}
```
#### Plot
```{r Mediation with latent fctrs late plot, eval=FALSE}
```

## Integrated/Quasi Mediation Analysis Code

### Early Integration of omics datasets

### Intermediate Integration of omics datasets

### Late Integration of omics datasets