# Supplemental Code {#sec-code}

## High Dimensional Mediation Analysis Code

```{r read-in_hima, include=FALSE}
knitr::read_chunk("project_setup/functions/hima_functions.R")
knitr::read_chunk("project_setup/functions/plot_hima_function.R")
```


### Early Integration of omics datasets {#sec-code_hima_early}

```{r HIMA_early, eval=FALSE}
```
### Intermediate Integration of omics datasets {#sec-code_hima_int}

```{r HIMA_intermediate, eval=FALSE}
```

### Late Integration of omics datasets {#sec-code_hima_late}

```{r HIMA_late, eval=FALSE}
```

### Plot HIMA {#sec-code_plot_hima}
```{r plot_HIMA, eval=FALSE}
```

## Mediation with Latent Factors Analysis Code

```{r read-in_mediation_with_latent_fctrs, include=FALSE}
knitr::read_chunk("project_setup/functions/mediation_with_latent_fctrs_functions.R")
knitr::read_chunk("project_setup/functions/plot_mediation_lf_function.R")
```

### Early Integration of omics datasets {#sec-code_med_with_latent_fctrs_early}

```{r Mediation_lf_early, eval=FALSE}
```

### Intermediate Integration of omics datasets {#sec-code_med_with_latent_fctrs_int}

```{r Mediation_lf_intermediate, eval=FALSE}
```

### Late Integration of omics datasets {#sec-code_med_with_latent_fctrs_late}

```{r Mediation_lf_late, eval=FALSE}
```

### Plot Mediation {#sec-code_plot_med_lf}
```{r plot_med_lf, eval=FALSE}
```

## Integrated/Quasi Mediation Analysis Code

```{r read-in_quasi_mediation, include=FALSE}
knitr::read_chunk("project_setup/functions/plot_omics_profiles.R")
knitr::read_chunk("project_setup/functions/plot_lucid_early.R")
knitr::read_chunk("project_setup/functions/plot_lucid_in_parallel.R")
knitr::read_chunk("project_setup/functions/plot_lucid_in_serial.R")
```

### Early Integration of omics datasets {#sec-code_plot_lucid_early}
```{r plot_LUCID_Early, eval=FALSE}
```

### Intermediate Integration of omics datasets {#sec-code_plot_lucid_int}
```{r plot_LUCID_in_Parallel, eval=FALSE}
```

### Late Integration of omics datasets {#sec-code_plot_lucid_late}
```{r plot_LUCID_in_Serial, eval=FALSE}
```

### Plot Omics Profile {#sec-code_plot_omics_profiles}
```{r plot_Omics_Profiles, eval=FALSE}
```

