# High Dimensional Mediation

## Set up project for High Dimensional Mediation Analysis

```{r 0. setup, include = FALSE, echo = FALSE}
source(fs::path(here::here(),"0_project_setup","libraries.R"))
source(fs::path(here::here(),"0_project_setup","directories.R"))
source(fs::path(here::here(),"0_project_setup","load_simu_data.R"))
source(fs::path(here::here(),"0_project_setup","functions","load_functions.R"))
source(fs::path(here::here(),
                "0_project_setup",
                "functions",
                "hima_functions.R"))

options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Early integration

```{r 1. early integration analysis}
# Run Analysis
result_hima_early <- hima_early_integration(exposure = exposure,
                                            outcome = outcome, 
                                            omics_lst = omics_lst, 
                                            covs = covs,
                                            Y.family = "gaussian", 
                                            M.family = "gaussian")

# Merge results with feature metadata 
result_hima_early_fin <- result_hima_early |>
  mutate(ftr_name = str_remove_all(ftr_name, "`")) |>
  tidylog::left_join(omics_names) |>
  mutate(across(where(is.character), ~na_if(., ""))) |>
  janitor::remove_empty(which = "cols") %>%
  mutate(omic_layer = ifelse(str_detect(ftr_name, "miR"), 
                             "miRNA",
                             omic_layer))

# Filter to significant features only and scale % total effect to 100
result_hima_early_fin <- result_hima_early_fin |> 
  tidylog::filter(BH.FDR < 0.05) |>
  mutate(pte_early = 100*`% total effect`/sum(`% total effect`))
```

## Intermediate Integration

**Steps:**

**- Regression 1 (linear regression): m_i = a0 + a1\*x, where x is exposure and m_i is each mediator.**

**- Regression 2 (group lasso): y = b0 + b1\*x + b2_i\*M, where y is outcome, x is exposure, and M is mediator matrix corresponding estimate b2_i**

**- Regression 3 (individual linear regressions): y = b0 + b1\*x + b2_star_i\*m_i, where y is outcome, x is exposure, m_i is individual mediator. se of b2_star_i used for calculating mediation confidence intervals**

**- Calculate mediation confidence interval for mediator i with:** - alpha = a1 - se of alpha = se of a1 - beta = b2_i - se of beta = se of b2_star_i,

```{r 2. intermediate integration analysis, cache=TRUE}
# # Run Analysis (takes ~45 minutes)
# system.time(
result_hima_intermediate <- hima_intermediate_integration(omics_lst = omics_lst,
                                                 covs = covs,
                                                 outcome = outcome,
                                                 exposure = exposure,
                                                 n_boot = 12,
                                                 Y.family = "gaussian")


# Merge results with feature metadata 
result_hima_intermediate_fin <- result_hima_intermediate |>
  mutate(ftr_name = str_remove_all(ftr_name, "`")) |>
  dplyr::select(-omic_layer) |>
  tidylog::left_join(omics_names) |>
  mutate(across(where(is.character), ~na_if(., ""))) |>
  janitor::remove_empty(which = "cols") %>%
  mutate(omic_layer = ifelse(str_detect(ftr_name, "miR"), 
                             "miRNA",
                             omic_layer))

# Filter to significant only and scale % total effect to 100
result_hima_intermediate_fin <- result_hima_intermediate_fin |> 
  tidylog::filter(sig_intermediate == 1) |>
  mutate(pte_intermediate = 100*pte_intermediate/sum(pte_intermediate))
```

## Late integration

```{r 3. late integration analysis}
# Run Analysis
result_hima_late <- hima_late_integration(exposure = exposure,
                                          outcome = outcome, 
                                          omics_lst = omics_lst, 
                                          covs = covs, 
                                          Y.family = "gaussian", 
                                          M.family = "gaussian")

# Merge results with feature metadata 
result_hima_late_fin <- result_hima_late |>
  mutate(ftr_name = str_remove_all(ftr_name, "`")) |>
  tidylog::left_join(omics_names) |>
  mutate(across(where(is.character), ~na_if(., ""))) |>
  janitor::remove_empty(which = "cols")

# Filter to significant features only, scale % total effect to 100
result_hima_late_fin <- result_hima_late_fin |> 
  tidylog::filter(BH.FDR < 0.05) |>
  mutate(pte_late = 100*`% total effect`/sum(`% total effect`))
```

## Figures: High dimensional mediation plots.

### Prepare data for figures.

```{r 4. Prepare Results for Figures}
early <- result_hima_early_fin |>
  tidylog::filter(BH.FDR < 0.05) |>
  mutate(sig_early = if_else(BH.FDR < 0.05, 1, 0)) |>
  select(ftr_name, 'alpha*beta', pte_early, alpha, beta, sig_early) |>
  rename(ie_early = 'alpha*beta', 
         alpha_early = alpha, 
         beta_early = beta)

late <- result_hima_late_fin |>
  tidylog::filter(BH.FDR < 0.05) |>
  mutate(sig_late = if_else(BH.FDR < 0.05, 1, 0)) |>
  select(ftr_name, 'alpha*beta', pte_late, alpha, beta, sig_late) |>
  rename(ie_late = 'alpha*beta', 
         alpha_late = alpha, 
         beta_late = beta)

# Filter out effect estimate of 0
intermediate <- result_hima_intermediate_fin |>
  tidylog::filter(s1 != 0) |>
  select(ftr_name, indirect, pte_intermediate, alpha, 
         beta_bootstrap,  sig_intermediate) |>
  rename(ie_intermediate = indirect, 
         alpha_intermediate = alpha, 
         beta_intermediate = beta_bootstrap)

# Combine all results
high_dim_res <- tidylog::full_join(early, late) |>
  tidylog::full_join(intermediate) |>
  mutate(omic_layer = get_omic_layer(ftr_name), 
         omic_layer = replace_na(omic_layer, "Metabolome"), 
         across(where(is.numeric), ~replace_na(., 0))) |>  
  rowwise() |>
  mutate(sum_pte = sum(c(pte_early + pte_late, pte_intermediate), 
                       na.rm = TRUE),
         sum_ie = sum(c(ie_early + ie_late, ie_intermediate), 
                      na.rm = TRUE),
         ftr_name = fct_reorder(ftr_name, sum_pte)) |>
  ungroup()

# Pivot data longer
high_dim_res_long <- high_dim_res |>
  pivot_longer(cols = c(ie_early:sig_intermediate), 
               names_to = c(".value", "method"), 
               names_sep = "_") |>
  arrange(ftr_name) |>
  mutate(method = str_to_title(method),
         omic_layer = as.factor(omic_layer) |>
           fct_relevel(c("Methylation", 
                         "Transcriptome", 
                         "miRNA", 
                         "Proteins")), 
         sig_chr = if_else(sig == 1, "*", ""), 
         ftr_name = rename_ftrs(ftr_name))

# Add names of features for plotting
high_dim_res_fin <- high_dim_res_long |>
  tidylog::left_join(omics_names |> dplyr::select(-omic_layer))
  # select(ftr_name_for_plots, everything())
```

### Figure. Correlation plot

```{r 5. correlation plot}
# Select the unique features from high dimensional mediation results
omics_df_subset <- omics_df %>%
  select(unique(high_dim_res_fin$ftr_name)) 
high_dim_res_fin$ftr_name[!(high_dim_res_fin$ftr_name %in% colnames(omics_df))]



# Create an annotation dataframe for the selected features
# Each feature name is annotated with its corresponding omic layer
annotation <- data.frame(
  ftr_name = colnames(omics_df_subset)) %>%
  mutate(Type = get_omic_layer(ftr_name))

# Rename columns of the omics dataframe using feature names meant for plots
setnames(omics_df_subset, old = colnames(omics_df_subset), 
         new = c(unique(high_dim_res_fin$ftr_name_for_plots)))

# Compute the correlation matrix for the omics dataframe
cormat <- cor(omics_df_subset)

# Create a heatmap of the correlation matrix using ComplexHeatmap package
# The heatmap splits rows and columns based on the annotation types 
p <- ComplexHeatmap::Heatmap(cormat, 
                             row_split = annotation$Type,
                             column_split = annotation$Type,
                             show_row_names = FALSE,
                             show_column_names = FALSE, 
                             heatmap_legend_param = list(title = "Correlation"))

# Extract the order of the features from the heatmap and combine into a single vector
feature_order <- c(ComplexHeatmap::row_order(p)$Methylation, 
                   ComplexHeatmap::row_order(p)$Transcriptome)

# Reorder feature names based on the feature order vector
feature_name_reorder <- colnames(omics_df_subset)[feature_order]

# Recalculate correlation matrix for the reordered dataframe
correlation_matrix <- omics_df_subset %>% 
  select(all_of(feature_name_reorder)) %>% 
  cor()

# Function to get the lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}

# Transform correlation matrix into a data frame, excluding NA values and redundant pairs
correlation_matrix_melt <- reshape2::melt(get_lower_tri(correlation_matrix), na.rm = TRUE)%>%
  filter(Var1 != Var2) %>%
  rename(correlation = value)

# Generate a plot using ggplot to visualize the correlation matrix
# The plot is a heatmap, with the fill color representing the correlation
p <- ggplot(correlation_matrix_melt,
            aes(x=Var1,y=Var2,fill=correlation)) + 
  geom_tile(aes(fill = correlation), colour = "white") +
  xlab(NULL)+
  ylab(NULL) +
  theme(
    axis.text.x=element_text(angle = 90, hjust = 1, size = 10),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.text.y=element_text(angle = 0, hjust = 1, size = 10),
    axis.ticks=element_blank(),
    panel.border=element_blank(),
    legend.position = "bottom",
    legend.key.width = unit(1, 'cm'),
    text = element_text(size = 12) ) +
  scale_y_discrete(position = "right") +
  scale_fill_gradient2(low  = "blue", mid  = "white", high = "red",
                       midpoint = 0, limits = c(-1, 1))

p

# Save the ggplot object to a file
ggsave(p, filename = fs::path(dir_figs,
                              "2_high_dim_multivar_mediation",
                              "correlaton_plot_070523.jpeg"),
       width = 9, height = 9.5, bg = "white")
```

## Figure 4. High dimensional mediation plots.

**High dimensional mediation with early (a), intermediate (b), and late (c) integration**

```{r 6. HIMA plot, fig.height=10}
# Early integration Plot
early_plot <- high_dim_res_fin %>% 
  filter(method == "Early") %>% 
  pivot_longer(cols = pte:beta) %>% 
  mutate(name = case_when(name == "alpha"~ "Alpha",
                          name == "beta" ~ "Beta",
                          name == "pte" ~ "TE (%)")) %>%
  ggplot(aes(x = factor(ftr_name_for_plots, 
                        levels = c(feature_name_reorder)), 
             y = value,
             fill = omic_layer, 
             alpha = sig_chr,
             label = sig_chr)) + 
  geom_bar(stat = "identity") +
  facet_grid(name ~ ., scales = "free", space = "free_x") +
  scale_fill_manual(values = c(sankey_colors$range[2], 
                               sankey_colors$range[5], 
                               sankey_colors$range[4], 
                               sankey_colors$range[3])) +
  scale_alpha_manual(values = c(.25, 1), guide = "none") + 
  geom_hline(yintercept = 0, linetype = 1, color = "grey50") + 
  ylab(NULL) + xlab(NULL) +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    axis.text.x = element_blank(),
    # axis.text.x = element_text(angle = 90, hjust = 1),
    legend.title = element_blank(), 
    legend.position = "none", # Place the legend at the bottom
    legend.box = "horizontal", # Align the legend horizontally
    legend.margin = margin(0, 0, 0, 0), # Remove extra margins around the legend
    legend.box.margin = margin(0, 0, 0, 0), # Adjust legend box margin if needed
    legend.direction = "horizontal", # Align the legend items horizontally
    legend.justification = c(1, 0))

# Late integration Plot
late_plot <- high_dim_res_fin %>% 
  filter(method == "Late") %>% 
  pivot_longer(cols = pte:beta) %>% 
  mutate(name = case_when(name == "alpha"~ "Alpha",
                          name == "beta" ~ "Beta",
                          name == "pte" ~ "TE (%)")) %>%
  ggplot(aes(x = factor(ftr_name_for_plots, levels = c(feature_name_reorder)), 
             y = value,
             fill = omic_layer, 
             alpha = sig_chr,
             label = sig_chr)) + 
  geom_bar(stat = "identity") +
  facet_grid(name ~ .,
             scales = "free",
             space = "free_x") + 
  scale_fill_manual(values = c(sankey_colors$range[2], 
                               sankey_colors$range[5], 
                               sankey_colors$range[4], 
                               sankey_colors$range[3])) +
  scale_alpha_manual(values = c(.25, 1), guide = "none") + 
  geom_hline(yintercept = 0, linetype = 1, color = "grey50") + 
  ylab(NULL) + xlab(NULL) +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    axis.text.x = element_text(angle = 90, hjust = 0),
    # axis.text.x = element_blank(),
    legend.title = element_blank(), 
    legend.position = "bottom", # Place the legend at the bottom
    legend.box = "horizontal", # Align the legend horizontally
    legend.margin = margin(0, 0, 0, 0), # Remove extra margins around the legend
    legend.box.margin = margin(0, 0, 0, 0), # Adjust legend box margin if needed
    legend.direction = "horizontal", # Align the legend items horizontally
    legend.justification = c(1, 0))

# Intermediate integration Plot
intermediate_plot <- high_dim_res_fin %>% 
  filter(method == "Intermediate") %>% 
  pivot_longer(cols = pte:beta) %>% 
  mutate(name = case_when(name == "alpha"~ "Alpha",
                          name == "beta" ~ "Beta",
                          name == "pte" ~ "TE (%)")) %>%
  ggplot(aes(x = factor(ftr_name_for_plots, levels = c(feature_name_reorder)), 
             y = value,
             fill = omic_layer,
             alpha = sig_chr,
             label = sig_chr)) + 
  geom_bar(stat = "identity") +
  facet_grid(name ~ ., scales = "free", space = "free_x") + 
  scale_fill_manual(values = c(sankey_colors$range[2], 
                               sankey_colors$range[5], 
                               sankey_colors$range[4], 
                               sankey_colors$range[3])) +
  scale_alpha_manual(values = c(.25, 1), guide = "none") + 
  geom_hline(yintercept = 0, linetype = 1, color = "grey50") + 
  ylab(NULL) + xlab(NULL) +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    axis.text.x = element_blank(),
    legend.title = element_blank(), 
    legend.position = "none", # Place the legend at the bottom
    legend.box = "horizontal", # Align the legend horizontally
    legend.margin = margin(0, 0, 0, 0), # Remove extra margins around the legend
    legend.box.margin = margin(0, 0, 0, 0), # Adjust legend box margin if needed
    legend.direction = "horizontal", # Align the legend items horizontally
    legend.justification = c(1, 0))

p <- plot_grid(NULL, early_plot, 
               NULL, intermediate_plot, 
               NULL, late_plot,
               ncol = 1,
               rel_heights = c(.03, 0.35, .03, 0.35, 0.03, 0.65),
               labels = c("a) Early", "","b) Intermediate","", "c) Late",""))
p

# Save Figure 
ggsave(p, filename = fs::path(dir_figs,
                              "2_high_dim_multivar_mediation",
                              "HIMA_plot_070523.jpeg"),
       width = 10, height = 12, bg = "white")
```
