# Introduction

## R Packages

```{r read_libraries, include=FALSE}
source(fs::path(here::here(),"project_setup","directories.R"))
knitr::read_chunk("project_setup/libraries.R")
```

The following is a list of packages that are used throughout this book that need to be loaded before any analysis. A complete list of all the packages used in the book, as well as code to download these packages, can be found in [Chapter @sec-software].
```{r load_packages, message = FALSE, warning=FALSE, collapse=FALSE}
#| code-fold: show
```

In order to replicate the style of the figures in this book, you will also have to set the ggplot theme:
```{r set_theme, message = FALSE, warning=FALSE}
#| code-fold: show
```

\newpage   

## The Data

```{r read-in load data, include=FALSE}
knitr::read_chunk("project_setup/load_simu_data.R")
```

The data used in this project is based off of simulated data from the Human Early Life Exposome (HELIX) cohort [@RN38]. The data was simulated for one exposure, five omics layers, and one continuous outcome (after publication, this data will be available on github). The format of this data is a named list with 6 elements. It includes separate numeric matrices for each of the 5 omics layers, as well as the exposure and phenotype data. In all datasets in the list, the rows represent individuals and the columns represent omics features. In this analysis, the exposure and outcome are:\

-  **Exposure**: *hs_hg_m_resid*, representing maternal mercury levels\
-  **Outcome**: *ck18_scaled*, representing child liver enzyme levels, a major risk factor for non-alcoholic fatty liver disease (NAFLD).  

This data is provided in the EnvirOmics package, which you can access in R using the following code:  

```{r load_data, collapse=FALSE} 
```


### Descriptive Statistics  

@tbl-descript shows the summary statistics for the exposure and phenotype data in this analysis.

```{r data summary}
#| tbl-cap: "Descriptive Statistics for the Simulated Variables"
#| label: tbl-descript
table1::table1(~., data = simulated_data[["phenotype"]][,-1])
```

\newpage   

### Mercury exposure and childhood MAFLD risk

```{r lr example, eval=FALSE}
lm_res <- lm(ck18_scaled ~ hs_hg_m_scaled + 
     e3_sex_None +
     hs_child_age_yrs_None,
   data = simulated_data[["phenotype"]])

summary(lm_res)
```

In the simulated data, each 1 standard deviation increase in maternal mercury was associated with a 0.11 standard deviation increase in CK18 enzymes (@fig-lr_res; p=0.02), after adjusting for child age and child sex. 

```{r linear regression}
#| fig-cap: "Association between maternal mercury and CK18 in the Simulated Data"
#| label: fig-lr_res

ggplot(data = simulated_data[["phenotype"]], 
       aes(x = hs_hg_m_scaled, y = ck18_scaled)) + 
  geom_point() +
  stat_smooth(method = "lm",
              formula = y ~ x ,
              geom = "smooth") +
  xlab("Maternal Mercury Exposure (Scaled)") +
  ylab("CK-18 Levels (Scaled)")
```

\newpage   

### Correlation of omics features

@fig-cor shows the correlation within and between the omics layers in the simulated data. 

```{r with ComplexHeatmap,fig.height= 7, fig.width= 6}
#| label: fig-cor
#| fig-cap: "Heatmap illustrating the correlation of molecular features within and between different omics layers." 
#| fig-height: 6

# Change omics list elements to dataframes
omics_df <- purrr::map(omics_lst, ~as_tibble(.x, rownames = "name")) %>%
  purrr::reduce(left_join, by = "name") %>%
  column_to_rownames("name")

meta_df <- imap_dfr(purrr::map(omics_lst, ~as_tibble(.x)),
                    ~tibble(omic_layer = .y, ftr_name = names(.x)))

# Correlation Matrix
cormat <- cor(omics_df, method = "pearson")

# Annotations
annotation <- data.frame(
  ftr_name = colnames(cormat),
  index = 1:ncol(cormat)) %>%
  left_join(meta_df, by = "ftr_name") %>%
  mutate(omic_layer = str_to_title(omic_layer))

# Make Plot
Heatmap(cormat, 
        row_split = annotation$omic_layer,
        column_split = annotation$omic_layer,
        show_row_names = FALSE,
        show_column_names = FALSE, 
        column_title_gp = gpar(fontsize = 12),
        row_title_gp = gpar(fontsize = 12),
        heatmap_legend_param = list(title = "Correlation"))
```


