# Mediation with latent factors {#sec-med_lf}

We define mediation with latent factors as a two step approach, in which we first perform dimensionality reduction on the omics data and then use the factors/clusters as latent mediator for the mediation analysis between the exposure and the outcome. 

## Set up project for High Dimensional Mediation Analysis
```{r set up data, include = TRUE, echo = TRUE, message=FALSE, warning=FALSE}
source(fs::path(here::here(),"project_setup","directories.R"))
source(fs::path(dir_proj, "libraries.R"))
source(fs::path(dir_proj, "load_simu_data.R"))
source(fs::path(dir_proj, "functions","mediation_with_latent_fctrs_functions.R"))
source(fs::path(dir_proj, "functions","plot_mediation_lf_function.R"))

options(knitr.table.format = "html")
```


## *Early integration*

For early integration, we used principal component analysis (PCA) as a dimensionality reduction step and selected the top i principal components which explained >80% of the variance. Following the joint dimensionality reduction step, we used the r package HIMA [@RN33] to examine whether the variance components mediated associations of in utero mercury exposure with MAFLD. The full code for the `med_lf_early` function is provided in @sec-code_med_with_latent_fctrs_early and `plot_med_lf` function is provided in @sec-code_plot_med_lf.

In this analysis,  principal components explained >80% of the variance in the combined omics datasets. Of these  components, 7 significantly mediated the relationship between maternal mercury and childhood liver injury (@fig-lf_early). 

### HIMA Early Integration
```{r 1a. early integration analysis, message=FALSE}
# Run Analysis
result_med_with_latent_fctrs_early <-
  med_lf_early(exposure, outcome,
               omics_lst, covs = NULL,
               Y.family = "gaussian",
               M.family = "gaussian",
               fdr.level = 0.05)
```

### Plot Early Integration
```{r 1b. plot early integration, message=FALSE}
#| label: fig-lf_early
#| fig-cap: "Mediation analysis with latent factors and early integration identifies joing components which mediate the association between maternal mercury and childhood liver injury. Panel A shows the mediation effects, where Alpha represents the coefficient estimates of the exposure to the mediator, Beta indicates the coefficient estimates of the mediators to the outcome, and TE (%) represents the percent total effect mediated calculated as alpha*beta/gamma. Panel B shows the individual correlation between the omic feature and the joint component." 
#| fig-width: 4
#| fig-height: 7

# Plot
(plot_med_lf(result_med_with_latent_fctrs_early))
```

## *Intermediate Integration*

The steps for intermediate integration start with performing a joint dimensionality reduction step using Joint and Individual Variance Explained (JIVE) [@RN18]. Following the joint dimensionality reduction step, we used the r package HIMA [@RN33] to examine whether the variance components mediated associations of in utero mercury exposure with MAFLD. 

### Conduct JIVE and Perform Mediation Analysis

#### Conduct JIVE
For this step, JIVE can estimate the optimal number of joint and individual ranks by changing the `method` argument in the function `jive`. For the simulated HELIX data, the optimal number, determined by setting `method = "perm"`, was 22 joint ranks and 6, 9, 5, 5, and 8 ranks for the methylome, transcriptome, miRNA, proteome, and metabolome, respectively.

#### Perform mediation analysis

In this analysis, 6 joint components, 1 transcriptome specific component significantly mediated the relationship between maternal mercury and childhood liver injury (@fig-lf_intermediate).The full code for the `med_lf_intermediate` function is provided in @sec-code_med_with_latent_fctrs_int and `plot_med_lf` function is provided in @sec-code_plot_med_lf. 

```{r JIVE mediation,message = FALSE, warning = FALSE}
# Run analysis
result_med_with_latent_fctrs_JIVE <- med_lf_intermediate(exposure, outcome,
                                          omics_lst, covs = NULL, 
                                          Y.family = "gaussian",
                                          M.family = "gaussian",
                                          fdr.level = 0.05)
```

### Plot Intermediate Integration

```{r plot JIVE mediation,message = FALSE, warning = FALSE}
#| label: fig-lf_intermediate
#| fig-cap: "Mediation analysis with latent factors and intermediate integration identifies joint and individual variance componets which mediate the association between maternal mercury and childhood liver injury. Panel A shows the mediation effects, where Alpha represents the coefficient estimates of the exposure to the mediator, Beta indicates the coefficient estimates of the mediators to the outcome, and TE (%) represents the percent total effect mediated calculated as alpha*beta/gamma. Panel B shows the individual correlation between the omic feature and the joint and individual components." 
#| fig-width: 4
#| fig-height: 7.2

(plot_med_lf(result_med_with_latent_fctrs_JIVE))
```

## *Late integration*

For late integration, we used principal component analysis (PCA) as a dimensionality reduction step on each omics layer separately, and selected the top i principal components which explained >80% of the variance. Following the dimensionality reduction step, we used the r package HIMA [@RN33]  to examine whether the variance components mediated associations of in utero mercury exposure with MAFLD. The full code for the `med_lf_late` function is provided in @sec-code_med_with_latent_fctrs_late and `plot_med_lf` function is provided in @sec-code_plot_med_lf

This analysis identified 2 methylated CpG sites, 1 miRNA, 1 protein and 2 expressed gene transcript clutesrs significantly mediated the association between mercury and MAFLD (@fig-lf_late).

### HIMA Late Integration
```{r pca late integration, message = FALSE, warning = FALSE}

result_med_with_latent_fctrs_late <-med_lf_late(exposure, outcome,
                                                omics_lst, covs = NULL,
                                                Y.family = "gaussian",
                                                M.family = "gaussian",
                                                fdr.level = 0.05)

```

### Plot Late Integration
```{r plot pca late integration, message = FALSE, warning = FALSE}
#| label: fig-lf_late
#| fig-cap: "Mediation analysis with latent factors and late integration identifies features in each omics layer individually which mediates the association between maternal mercury and childhood liver injury. Panel A shows the mediation effects, where Alpha represents the coefficient estimates of the exposure to the mediator, Beta indicates the coefficient estimates of the mediators to the outcome, and TE (%) represents the percent total effect mediated calculated as alpha*beta/gamma. Panel B shows the individual correlation between the omic feature and components." 
#| fig-width: 4
#| fig-height: 7

(plot_med_lf(result_med_with_latent_fctrs_late))
```

## Pathway Analysis.

Following mediation analysis, you can use the correlation p-values to perform pathway analysis with appropriate pathway analysis software.
