# Helix Project

```{r overall setup, include=FALSE}
source(here::here("0_project_setup/libraries.R"))
source(here::here("0_project_setup/directories.R"))
source(here::here("0_project_setup/load_data.R"))
source(here::here("0_project_setup/functions/load_functions.R"))

options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Description

In this analysis, the exposure and outcome are: 
Exposure: Maternal Mercury (hs_hg_m_resid)
Outcome: Liver Injury (ck18_scaled)


# 0. Project Setup
```{r child="0_project_setup/Project Setup with Mercury.Rmd"}

```


# 1. Descriptive Statistics
## Correlation of omics features
```{r child="1_descriptive_statistics/correlation of omics features.Rmd"}

```


## Data Summary
```{r child="1_descriptive_statistics/data summary.Rmd"}

```


## Exposure Outcome Associations
```{r child = "1_descriptive_statistics/exposure outcome associations.Rmd"}

```


# 2. High Dimensional mediation analysis
```{r child="2_high_dim_multivar_mediation/2_HIMA.Rmd"}

```

# 3. High Dimensional mediation analysis with Latent factors
```{r child="3_high_dim_multivar_mediation_latent_fctrs/High_dim_mediation_latent_fctrs.Rmd"}

```

# 4. Integrated latent unknown clusters
## Lucid early integration
```{r child="4_integrated_latent_unknown_clusters/Lucid Early Integration.Rmd"}

```

## Lucid in Parallel
```{r child="4_integrated_latent_unknown_clusters/Lucid in Parallel.Rmd"}

```

## Lucid in Serial
```{r child="4_integrated_latent_unknown_clusters/Lucid in Serial.Rmd"}

```
