# Software {#sec-software}

## *Installing R packages*

```{r read_libraries, include=FALSE}
source(fs::path(here::here(),"project_setup","directories.R"))
knitr::read_chunk("project_setup/libraries.R")
```

The following code will download all the packages needed to run the analyses in this book.
```{r install_packages, message = FALSE, warning=FALSE, collapse=FALSE}
#| code-fold: show
```


## *R development environment*

The following table lists the package versions and corresponding repository of all the R packages that were used to develop the code in this book. 

\renewcommand{\arraystretch}{0.8} <!-- decrease line spacing for the table -->
```{r packages, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
library(devtools)
source(fs::path(here::here(),"project_setup","directories.R"))
source(fs::path(dir_proj,"libraries.R"))

# get session info
df_packages <- devtools::session_info()$packages %>% 
  as.data.frame(.) %>% 
  remove_rownames(.) %>% 
  mutate(attached = if_else(attached == TRUE, "Attached", "Loaded")) %>%
  arrange(attached) %>% 
  select(package, loadedversion, attached, date, source)

colnames(df_packages) <- c("Package", "Version", "Attached or Loaded",
                           "Date/Publication", "Source")

kableExtra::kable(df_packages, 
                  booktabs = T, 
                  align = "l", 
                  longtable=T, 
                  caption = "List of all R Packages used in this book.") %>% 
  kableExtra::kable_styling(full_width = F,
                            latex_options = c("repeat_header", "striped"), 
                            position = "center")


```
\renewcommand{\arraystretch}{1} <!-- reset row height/line spacing -->

## *Software Environment* 

This book was developed on the following platform. 

\renewcommand{\arraystretch}{0.8} <!-- decrease line spacing for the table -->
```{r Reproducibility-SessionInfo-R-environment, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}

# get session info
df_session_platform <- devtools::session_info()$platform %>% 
  unlist(.) %>% 
  as.data.frame(.) %>% 
  rownames_to_column(.)

colnames(df_session_platform) <- c("Setting", "Value")

kableExtra::kable(
  df_session_platform, 
  booktabs = T, 
  align = "l") %>% 
  kableExtra::kable_styling(full_width = F,
                            latex_options = c("repeat_header", "striped"), 
                            position = "center")
```
\renewcommand{\arraystretch}{1} <!-- reset row height/line spacing -->